---
title: 'Respiratory Viruses'
output: word_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = NA)

knitr::opts_chunk$set(
    comment = "",
    echo = F,
    fig.width = 12, 
    fig.height = 10,
    message = FALSE,
    warning = FALSE,
    tidy.opts = list(
        keep.blank.line = TRUE,
        width.cutoff = 150
        ),
    options(width = 150),
    eval = TRUE
)

#this chunk is used to silence code, warnings, comments and hash from appearing in output

# useful links 
# r - Split long gtsummary() table to n smaller tables - Stack Overflow
# https://stackoverflow.com/questions/66119997/split-long-gtsummary-table-to-n-smaller-tables

```


```{r import and cleaning}
library(tidyverse)
library(readxl)
library(gtsummary)
library(labelled)

data <- read_excel("R.viral screening-final-EXCELED-3.21.xlsx")
names(data) <- make.names(names(data), unique = T)


```



```{r functions}

rmdtbl <- function(df){
  
  tbl_alpha <- autofit(theme_vanilla(flextable(df)))
  
  tbl_alpha <- bg(tbl_alpha, bg = "blue", part = "header")
  tbl_alpha <- color(tbl_alpha, color = "white", part = "header")
  
  
  bes <- align(tbl_alpha, align = "center")
  
  bes <- align_text_col(bes, align = "center")
  return(bes)
}
```


```{r cleaning}

data$pre.transplant.R..viral.screening <-   fct_collapse(data$pre.transplant.R..viral.screening, 
Negative = "negative", 
other_level = "Positive")

data$Recipient.s.Gender. <- fct_collapse(data$Recipient.s.Gender., 
                                         Male = "Male", 
                                         other_level = "Female")

data$Stem.cell.source <- tolower(data$Stem.cell.source)

data <- data %>% 
  mutate_if(is.character, list(~na_if(., "_"))) # converting "_" into NA

data$type.of.R.viruses.if....ve..AdV <- fct_collapse(data$type.of.R.viruses.if....ve..AdV,
                                                     AdV = "Adv")

data$Age <- (data$Date.of.Stem.Cell.Infusion. - data$Recipient.s.Birth.Date.)/365.25 

```

# Patients characteristics

```{r table1}

var_label(data) <- list(
pre.transplant.R..viral.screening = "Viral screening", 
Recipient.s.Gender. = "Recipient gender", 
Underlying.malignancy = "Diagnosis",
Type.of.transplant = "Donor type",
Stem.cell.source = "stem cell source", 
conditioning.Regimen = "conditioning regimen", 
Pre.transplant.chest.morbidity = "chest morbidity",
GVHD.ppx..CSA.MTX = "GVHD prophylaxis[CSA+MTX]", 
initial.neutrophils.count.on.admission. = "Neutrophil count", 
Initial.lymphocytes.count.on.admission. = "Lymphocyte count", 
T.cell.subset.CD19.interpretation. = "CD19 count", 
T.cell.subset.CD3.interpretation. = "CD3 count", 
T.cell.subset.CD56..CD16.interpretation. = "CD56 + CD16 counts", 
type.of.R.viruses.if....ve.HRV = "HRV", 
type.of.R.viruses.if....ve..AdV = "AdV",
type.of.R.viruses.if....ve..others = "Other viruses")

tbl1 <- data %>% 
  dplyr::select(pre.transplant.R..viral.screening, Recipient.s.Gender., Age, Type.of.transplant, Stem.cell.source, conditioning.Regimen,
                Pre.transplant.chest.morbidity, GVHD.ppx..CSA.MTX, initial.neutrophils.count.on.admission., Initial.lymphocytes.count.on.admission., T.cell.subset.CD19.interpretation., T.cell.subset.CD3.interpretation., 
T.cell.subset.CD56..CD16.interpretation.)

tbl_summary(tbl1, by = pre.transplant.R..viral.screening, missing = "no", sort = list(everything() ~ "frequency")) %>% add_p() %>% add_overall()  %>% bold_labels() %>% italicize_levels() %>% bold_p() %>% gtsummary::as_flex_table() 

```

Table1 patients' characteristics grouped by pre-transplant viral screening
\n

N.B.
chest morbidity is pre-transplant \n
counts are on admission 

# Types of viruses


```{r}
data <- data %>% 
  unite("viruses", type.of.R.viruses.if....ve.HRV : type.of.R.viruses.if....ve..others, sep = "/", remove = FALSE, na.rm = T)

data$viruses[data$viruses == ""] <- NA

tbl2 <- data %>% 
  dplyr::select(viruses)

tbl_summary(tbl2, missing = "no", sort = list(everything() ~ "frequency"), percent = "cell")  %>% add_n() %>% bold_labels() %>% italicize_levels() %>% gtsummary::as_flex_table() 

```
Table 2 Types of viruses

N.B. AdV in this case means patients that had only AdV positive test. 


# Types of viruses after grouping 

```{r table 3}

data$other_viruses <- ifelse(is.na(data$type.of.R.viruses.if....ve..others), "no", "yes")
data$type.of.R.viruses.if....ve.HRV <- ifelse(is.na(data$type.of.R.viruses.if....ve.HRV), "no", "yes")
data$type.of.R.viruses.if....ve..AdV <- ifelse(is.na(data$type.of.R.viruses.if....ve..AdV), "no", "yes")

var_label(data) <- list(
type.of.R.viruses.if....ve.HRV = "HRV", 
type.of.R.viruses.if....ve..AdV = "AdV",
type.of.R.viruses.if....ve..others = "Other viruses")


tbl3 <- data %>% 
  dplyr::select(type.of.R.viruses.if....ve..AdV,  type.of.R.viruses.if....ve.HRV, 
                other_viruses)

tbl_summary(tbl3, missing = "no", sort = list(everything() ~ "frequency")) %>% bold_labels() %>% italicize_levels() %>% gtsummary::as_flex_table() 

```

Table 3: Number of patients having AdV or HRV, regardless of other concomitant viruses. 

